---
title: "Rcpp nanoflann"
author: "Thell"
date: "12/16/2014"
output: html_document
---

Timing and output information in response to [SOQ-27321856](http://stackoverflow.com/questions/27321856/closest-point-to-a-path).

__using path and centers *values* as defined in the orginal question__

````{r setup, echo=FALSE}
set.seed(1)
n <- 10000
x <- 100*cumprod(1 + rnorm(n, 0.0001, 0.002))
y <- 50*cumprod(1 + rnorm(n, 0.0001, 0.002))

path <- data.frame(cbind(x=x, y=y))

centers <- expand.grid(x=seq(0, 500,by=0.5) + rnorm(1001), 
                       y=seq(0, 500, by=0.2) + rnorm(2501))
````

To achieve identical results to the original the RANN solution needs
slight modification which we time here...

````{r}
library(RANN)
system.time(o.flann<-unique(as.numeric(nn2(centers,path,1)$nn.idx)))
````

The Rcppnanoflann solution takes advantage of Rcpp, RcppEigen and
the nanoflann EigenMatrixAdaptor along with the c++11 <set> to yield
identical unique indexes to the original question.
````{r}
library(Rcppnanoflann)
system.time(o.nano<-nnIndex(centers,path))
````

````{r}
identical(o.flann,o.nano)
````

The working function of Rcppnanoflann takes advantage of Eigen's Map
capabilities to create the input for a fixed type Eigen matrix from
the given `P` dataframe.

I tested the RcppParallel package but the kd_tree object does not have
a copy constructor, so the tree needed to be created for each thread
which ate up any gains in the parallel query processing.

RcppEigen and Rcpp11 currently don't play together so the idea of
using Rcpp11's parallel sapply for the query isn't easily tested.

````{r,eval=FALSE,tidy=TRUE}
// [[Rcpp::export]]
std::vector<double> nnIndex(const Rcpp::DataFrame & P, const Rcpp::DataFrame & Q )
{
  using namespace Eigen;
  using namespace Rcpp;
  using namespace nanoflann;

  // Matrix of points to be queried against.
  const NumericVector & Px(P[0]);
  const NumericVector & Py(P[1]);
  MatrixX2d M(Px.size(), 2);
  M.col(0) = VectorXd::Map(&Px[0],Px.size());
  M.col(1) = VectorXd::Map(&Py[0],Py.size());

  // The points to query.
  const NumericVector & Qx(Q[0]);
  const NumericVector & Qy(Q[1]);
  double query_pt[2];
  size_t query_count(Qx.size());

  // Populate a 2d tree.
  KD_Tree kd_tree( 2, M, 10 );
  kd_tree.index->buildIndex();

  std::set<size_t> nn;
  std::vector<double> out;
  out.reserve(query_count);

  size_t index(0);
  double quadrance;
  for( size_t i=0 ; i < query_count; ++i ) {
    query_pt[0] = Qx[i];
    query_pt[1] = Qy[i];
    kd_tree.index->knnSearch( &query_pt[0],1, &index, &quadrance);
    if( nn.emplace(index).second ) out.emplace_back(index+1);
  }

  return out;
}
````
